Michael Rojas <br>
U49602685 <br>
MA415/615 <br>
Feb 21, 2018


# 10.5 Exercises
## Q1. 
How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).

```{r}
mtcars
class(mtcars)
class(as_tibble(mtcars))
mt_tib <- as_tibble(mtcars)
mt_tib
```


## Ans1
Tibbles will only print out a limited number of rows and show the class on top of each column

##########

## Q2.
Compare and contrast the following operations on a data.frame and equivalent tibble. 
What is different? Why might the default data frame behaviors cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]

df$y
df$a
```



## Ans2
When Using a \$ a data.frame will partially complete the column. So even though we wrote "df\$x" it returned 
the value for df\$xyz. Dont know if this might be by mistake on r's part but it might bring up errors if 
there was a variable x. It can be seen above that i tried returning the values by substituting the value 
for "df$y" but it returned NULL values.
With data.frames, with [ the type of object that is returned differs on the number of columns. 
If it is one column, it won't return a data.frame, but instead will return a vector. With more than 
one column, then it will return a data.frame. 

## 3. 
If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the
reference variable from a tibble?

## Ans3

You can use the double bracket, like df[[var]]. You cannot use the dollar sign, because df$var would 
look for a column named var.

# Q4.
Practice referring to non-syntactic names in the following data frame by:

* i. Extracting the variable called 1.

* ii. Plotting a scatterplot of 1 vs 2.

* iii. Creating a new column called 3 which is 2 divided by 1.

* iv. Renaming the columns to one, two and three.

## Ans4 i.
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```
In order to extract the variable "a" we can do the following:
```{r}
annoying[["1"]]
```

## Ans4 ii.
A scatter plot of 1 vs. 2:

```{r}
ggplot(annoying, aes(x = `1`, y = `2`)) + geom_point()
```

## Ans4 iii.
A new column 3 which is 2 divided by 1:
```{r}
annoying[["3"]] <- annoying[["2"]] / annoying[["1"]]
annoying
```

## Ans4 iv.
Renaming the columns to one, two, and three:
```{r}
annoying <- rename(annoying, one = `1`, two = `2`, three = `3`)
glimpse(annoying)
```

## Q5
What does tibble::enframe() do? When might you use it?
```{r}
?tibble::enframe()
var <- c(a = 1, b = 2, c = 3)
var
enframe(c(a = 1, b = 2, c = 3))
```
 
## Ans5
It converts named vectors to a data frame with names and values makes it look cleaner as well.
Tidys things up in a nice format. Can be used to tidy up a vector.

## Q6
What option controls how many additional column names are printed at the footer of a tibble?

## Ans6
The print function for tibbles is in print.tbl_df:
Once looking at the help function for this it is seen that the n_extra takes care of this for us.
The option n_extra determines the number of extra columns to print abbreviated information for, if the width is too small for the entire tibble.If NULL, the default, will print information about at most tibble.max_extra_cols extra columns.
```{r}
?print.tbl_df
```

################################################
################################################

# 12.6.1 Exercises

## Q1
In this case study I set na.rm = TRUE just to make it easier to check that we had the correct values. Is this reasonable? Think about how missing values are represented in this dataset. Are there implicit missing values? What’s the difference between an NA and zero?


To start the following code is included as it might be of use for the questions:
```{r}
who1 <- who %>%
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
glimpse(who1)

who2 <- who1 %>%
 mutate(key = stringr::str_replace(key, "newrel", "new_rel"))

who3 <- who2 %>%
  separate(key, c("new", "type", "sexage"), sep = "_")
who3

who3 %>%
  count(new)

who4 <- who3 %>%
  select(-new, -iso2, -iso3)

who5 <- who4 %>%
  separate(sexage, c("sex", "age"), sep = 1)
who5
```

## Ans1
I think I would need to know more about the datas generation process. There are 0's in the data, which might depict that there are no cases for this fot the particular part of the data. We can thus filter these cases and pipe in into nrow() as follows:

```{r}
who1 %>%
  filter(cases == 0) %>% nrow()
```

So it appears that either a country has all its values in a year as non-missing if the WHO collected data for that country, or all its values are non-missing. So it is okay to treat explicitly and implicitly missing values the same, and we don't lose any information by dropping them.

```{r}
  gather(who, new_sp_m014:newrel_f65, key = "key", value = "cases") %>%
  group_by(country, year)  %>%
  mutate(missing = is.na(cases)) %>%
  select(country, year, missing) %>%
  distinct() %>%
  group_by(country, year) %>%
  filter(n() > 1)

```


## Q2
What happens if you neglect the mutate() step? (mutate(key = stringr::str_replace(key, "newrel", "new_rel")))

## Ans2

```{r}
who3a <- who1 %>%  separate(key, c("new", "type", "sexage"), sep = "_")

filter(who3a, new == "newrel") %>% head()
```

The separate function emits the warning "too few values", and if we check the rows for keys beginning with "newrel_", we see that sexage is messing, and type = m014.

## Q3
I claimed that iso2 and iso3 were redundant with country. Confirm this claim.

## Ans3
```{r}
select(who3, country, iso2, iso3) %>% distinct() %>% group_by(country) %>% filter(n() > 1)
```

## Q4
For each country, year, and sex compute the total number of cases of TB. Make an informative visualization of the data.

## Ans4
```{r}
who5 %>%
  group_by(country, year, sex) %>%
  filter(year > 1995) %>%
  summarise(cases = sum(cases)) %>%
  unite(country_sex, country, sex, remove = FALSE) %>%
  ggplot(aes(x = year, y = cases, group = country_sex, colour = sex)) +
  geom_line()
```

